Hybrid Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-08-12 manuscript/main.tex
lemma 4.4 revised, no regularization

\begin{frontmatter}
\title{ Partial Least Squares Regression with Multiple Functional and Scalar Predictors}
%\title{A sample article title with some additional note\thanksref{t1}}
\runtitle{A sample running head title}
%\thankstext{T1}{A sample additional note to the title.}

@@

\subsection{Preliminary step 1: finite-basis approximation}\label{sec: finite basis approximation}	
Let ${b_m(t)}$ be a twice-differentiable basis of $L^2([0,1])$ whose second derivatives are also linearly independent, for example, cubic B-splines, the Fourier basis, or an orthonormal polynomial basis of degree greater than three. Using this basis, 
the $j$th functional predictor, 
regression coefficient, PLS component direction, and orthogonalization regression coefficient (with iteration indices suppressed) are represented as follows:
\begin{equation*}
	X_{ij}(t) = \sum_{m=1}^\infty \theta_{ijm} \, b_m(t),~
	\beta_j(t) = \sum_{m=1}^\infty \eta_{jm} \, b_m(t),~
	%%%%%%%%%%%%%%%%%%%%%%
	\xi_j(t)
	=
	\sum_{m=1}^\infty
	\gamma_{jm}
	b_m(t),
	~
	\delta_j(t)
	=
	\sum_{m=1}^\infty
	\pi_{jm}
	b_m(t)		
\end{equation*}	
In practice, the full set of coefficients can not be obtained  with finite sample size, as functional data are measured on a finite grid. Thus, we truncate the expansion at $M$ terms and define the truncated basis space
$
\mathcal{B} := \operatorname{span}\bigl(b_1(t), \ldots, b_M(t)\bigr).
$
We choose a moderately large $M$ (e.g., 15 or 20) to capture functional variation without fine-tuning, as smoothness is handled via penalization (see Section~\ref{sec: Regularization of PLS Components}).
The truncated expansions of the predictor and coefficient are denoted as
\begin{equation*}
	\widetilde{X}_{ij}(t) := \sum_{m=1}^M \theta_{ijm} \, b_m(t), \quad
	\tilde{\beta}_j(t) := \sum_{m=1}^M \eta_{jm} \, b_m(t),
\end{equation*}
and our suggest method restricts each PLS component direction and orthogonalization regression coefficient to admit the following expansion:
\begin{equation}\label{def:truncated_zeta_and_delta}
	\xi_j(t) = \sum_{m=1}^M \gamma_{jm} \, b_m(t),
	\quad
	\delta_j(t)
	=
	\sum_{m=1}^M
	\pi_{jm}
	b_m(t)		
\end{equation}
This implies that all computations in this paper are carried out entirely within the subspace
\begin{equation}\label{def:truncated_hilbert_space}
	\widetilde{ \mathcal{H} } := \mathcal{B}^K \times \mathbb{R}^p \subset \mathcal{H}.
\end{equation}
The $i$th hybrid predictor, projected on $\widetilde{ \mathcal{H} }$, is represented by the tuple
\begin{equation*}
	\widetilde{W}_i := 
	(
	\widetilde{X}_{i1}, \ldots, \widetilde{X}_{iK}, \boldsymbol{Z}_i
	).
\end{equation*}
Let $\boldsymbol{\theta}_{ij}$, $\boldsymbol{\eta}_j$,  $\boldsymbol{\gamma}_j$, and
$\boldsymbol{\pi}_j$
denote the $M$-dimensional vectors of coefficients:
\begin{equation}\label{def:coef_vector_form}
	\boldsymbol{\theta}_{ij} := (\theta_{ij1}, \ldots, \theta_{ijM})^\top,
	~
	\boldsymbol{\eta}_{j} := (\eta_{j1}, \ldots, \eta_{jM})^\top ,~
	\boldsymbol{\gamma}_{j} := (\gamma_{j1}, \ldots, \gamma_{jM})^\top,~
	\boldsymbol{\pi}_{j} := (\pi_{j1}, \ldots, \pi_{jM})^\top
\end{equation}
For the predictors, we stack the coefficient vectors across observations into the matrix
\begin{equation}\label{def:theta}
	\Theta_j := (\boldsymbol{\theta}_{1j}, \ldots, \boldsymbol{\theta}_{nj})^\top  \in \mathbb{R}^{n \times M},
\end{equation}
and construct the full design matrix 
\begin{equation}\label{def:full_theta}
\Theta := (\Theta_1, \ldots, \Theta_K, Z) \in \mathbb{R}^{n \times (MK + p)}.
\end{equation}
Let us denote the response vector as $\mathbf{y} := (y_1, \ldots, y_n)^\top$.
Let $B, B^{\prime\prime} \in \mathbb{R}^{M \times M}$ denote the Gram matrices of the basis functions and their second derivatives, with entries
\begin{equation}\label{def:B_matrix}
	B_{m, m'} := \int_0^1 b_m(t)\, b_{m'}(t)\, dt, \quad
	B^{\prime\prime}_{m, m'} := \int_0^1 b_m''(t)\, b_{m'}''(t)\, dt,
\end{equation}
for $m, m' = 1, \ldots, M$. We then define the block-diagonal matrices
\begin{equation}\label{def:gram_block}
	\mathbb{B} := \operatorname{blkdiag}(B, \ldots, B, I_p), \quad
	\mathbb{B}^{\prime\prime} := \operatorname{blkdiag}(B^{\prime\prime}, \ldots, B^{\prime\prime}, I_p),
\end{equation}

@@

\subsubsection{Iterative step 2: residualization via hybrid-on-scalar regression} \label{section:sub:residualization}
The $l$-th iteration's second step involves residualization of both predictors and responses. We first compute the individual PLS score:  
\begin{equation}\label{def:plsscore}
	\hat{\rho}^{[l]}_i := \langle \widetilde{W}_i^{[l]}, \, \hat{\xi}^{[l]} \rangle_{\mathcal{H}},
\end{equation}
using the estimated PLS component direction $\widehat{\xi}^{[l]}$ obtained from Propositions \ref{proposition:eigen_regul} and \ref{proposition:linear_regul}. 
Since   $\widetilde{W}_i^{[l]}$ are assumed to have a sample mean of zero, these PLS scores will also have a sample mean of zero.
Using these PLS scores as scalar predictor observations, the $(l+1)$-th iteration's residualized response is computed as:
\begin{equation}\label{algorithm_step:residualization_response}
	Y_i^{[l+1]} := Y_i^{[l]} - 
	\frac{
		( \mathbf{y}^{[l]} )^\top  \hat{\boldsymbol{\rho}}^{[l]}
	}{
		( \hat{\boldsymbol{\rho}}^{[l]} )^\top  \hat{\boldsymbol{\rho}}^{[l]}
	}
	\widehat{\rho}_{il},
\end{equation}
which is equivalent to using the least squares estimate from regressing $Y_i^{[l]}$ on $\widehat{\rho}_{il}$,
a procedure identical to that found in the naive PLS algorithm (Section \ref{sec: Naive PLS Algorithm}). 
Since $Y_i^{[l]}$ are assumed to have a sample mean of zero, these residuals $Y_i^{[l+1]}$   also have a sample mean of zero.

For predictor residualization, we fit a hybrid-on-scalar regression model where the predictor is the PLS score for each observation. Specifically, we fit $\widetilde{W}^{[l]}_i = \hat{\rho}^{[l]}_i\, \delta^{[l]} + E_i^{[l]}$, and then take the residuals as the predictor for the next iteration.
Here, $\delta^{[l]}$ and $E_i^{[l]}$ are the regression coefficient and error terms, respectively.
Performing this pointwisely, as in Section \ref{sec: Naive PLS Algorithm}, is highly inefficient and potentially infeasible for dense or irregular functional data. Moreover, the results may lack smoothness, introducing instability that propagates to subsequent results and compromises the entire algorithm.
Therefore, we employ a basis expansion approach to fit the entire hybrid object in one step. We assume both $\delta^{[l]}$ and $E_i^{[l]}$ are elements of $\tilde{\mathcal{H}}$, as defined in \eqref{def:truncated_hilbert_space}. %To ensure the stability of the entire PLS estimation procedure, the estimated regression coefficient $\hat{\delta}^{[l]} = (\hat{\delta}^{[l]}_1, \ldots, \hat{\delta}^{[l]}_K, \hat{\boldsymbol{\chi}}^{[l]} ) \in \tilde{\mathcal{H}}$ should have smooth functional parts, consistent with Section \ref{sec: Regularization of PLS Components}.
 Consequently, $\delta^{[l]}$ is obtained by minimizing a %penalized 
 least squares criterion:%, where the functional part is penalized by the integrated squared second derivative defined in \eqref{def:roughness_penalty}:
 \begin{equation}\label{def:argmin_pensse}
 	\hat{\delta}^{[l]} := \arg \min_{\delta \in \tilde{\mathcal{H}}} \sum_{i=1}^n 
 	\| \widetilde{W}_i^{[l]} - \hat{\rho}_i^{[l]}\, \delta\|_{\mathcal{H}}^2.
 \end{equation}
 \begin{comment}
\begin{equation}\label{def:argmin_pensse}
	\hat{\delta}^{[l]} := \arg \min_{\delta \in \tilde{\mathcal{H}}} \operatorname{PENSSE}( \delta),
\end{equation}
where
\begin{equation}\label{def:pensse}
	\operatorname{PENSSE}( \delta)
	:=
	\sum_{i=1}^n \langle \widetilde{W}_i^{[l]} - \hat{\rho}_i^{[l]}\, \delta,
	\,
	\widetilde{W}_i^{[l]} - \hat{\rho}_i^{[l]}\, \delta
	\rangle_{\mathcal{H}}
	+
	\tau
	\sum_{j=1}^K\operatorname{PEN}(\delta_j).
\end{equation}
where $\tau>0$ is a  smoothing parameter controlling the trade-off between fit accuracy and functional regression coefficient smoothness.
\end{comment}
The following proposition shows that the solution of the problem \ref{def:argmin_pensse} can be obtained by solving linear equations.

\begin{lemma}[Closed-form  solution]\label{proposition:closed_form_orthgonalization}
	Let us denote
	$
	\hat{\boldsymbol{\rho}}^{[l]} := (\hat{\rho}^{[l]}_1, \ldots, \hat{\rho}^{[l]}_n)^\top.
	%\quad\text{and} \quad		\bar{\rho}^{[l]} := (\hat{\boldsymbol{\rho}}^{[l]})^\top \hat{\boldsymbol{\rho}}^{[l]}.
	$
	The solution to the optimization problem \eqref{def:argmin_pensse}, denoted
	$\hat{\delta}^{[l]}   \in \tilde{\mathcal{H}}$, has basis coefficients
	\begin{equation*}
		\hat{\boldsymbol{\pi}}_j^{[l]}
		=
		\frac{1}{\| 	\hat{\boldsymbol{\rho}}^{[l]}\|_2^2}
		\Theta_j^{[l] \top}
		\hat{\boldsymbol{\rho}}^{[l]}
		,
		~
		j = 1, \ldots, K,
	\end{equation*}
	  and scalar parts given by:
	 \begin{equation*}
	 	\hat{\boldsymbol{\chi} }^{[l]}
	 	=
	\frac{1}{\| 	\hat{\boldsymbol{\rho}}^{[l]}\|_2^2}
		\mathbf{Z}^{[l] \top}  \hat{\boldsymbol{\rho}}^{[l]}.
	 \end{equation*}
\end{lemma}
Proof of Lemma \ref{proposition:closed_form_orthgonalization} is provided in Appendix \ref{section:proof:proposition:closed_form_orthgonalization}.
\jongmin{proof needs to be revised}
The $(l+1)$-th iteration's residualized predictor is then computed as:
\begin{equation}\label{algorithm_step:residualization_predictor}
	\widetilde{W}_i^{[l+1]} 
	= 
	\widetilde{W}_i^{[l]}  - \widehat{\rho}_{il}  \widehat{ {\delta}}_{l}.
	%%%%%%%%%%%
\end{equation}
Since $\widetilde{W}_i^{[l]} $ are assumed to have a sample mean of zero, these residuals $\widetilde{W}_i^{[l+1]}$   also have a sample mean of zero.

@@

\bibliographystyle{apalike}
\bibliography{Reference}

@@

@@

Now to switch our gears toward the the orthogonality. For  $l_1 > l_2 $, we have
\begin{align*}
	\langle \widehat{\xi}_{l_1}, \widehat{\xi}_{l_2} \rangle_{\mathcal{H}, \Lambda}
	& \overset{(i)}{=}
	\hat{\boldsymbol{\xi}}_{l_1}^\top 
	\hspace{-.2em}
	(\mathbb{B} + \Lambda \mathbb{B}^{\prime \prime}) \hat{\boldsymbol{\xi}}_{l_2}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\\	& \overset{(ii)}{=} \frac{1}{\kappa_{l_1}} (V^{[l_1]} \widehat{\boldsymbol{\xi}}_{l_1})^\top \hat{\boldsymbol{\xi}}_{l_2}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\\	& = \frac{1}{\kappa_{l_1}}
	\widehat{\boldsymbol{\xi}}_{l_1}^\top
	V^{[l_1]} \hat{\boldsymbol{\xi}}_{l_2} 
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\\	& \overset{(iii)}{=} \frac{1}{ n^2 \kappa_{l_1}}
	\widehat{\boldsymbol{\xi}}_{l_1}^\top
	(\mathbb{B} \Theta^{[l_1]\top} \mathbf{y})(\mathbb{B} \Theta^{[l_1]\top} \mathbf{y})^\top
	\hat{\boldsymbol{\xi}}_{l_2}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\\	& = \frac{1}{ n^2 \kappa_{l_1}}
	\widehat{\boldsymbol{\xi}}_{l_1}^\top
	(\mathbb{B} \Theta^{[l_1]\top} \mathbf{y})
	\mathbf{y}^\top
	\Theta^{[l_1] } 
	\mathbb{B}
	\hat{\boldsymbol{\xi}}_{l_2}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\\	& = c
	\mathbf{y}^\top
	(
	\Theta^{[l_1] }
	\mathbb{B}
	\hat{\boldsymbol{\xi}}_{l_2}
	),
\end{align*}
where
step $(i)$ uses \eqref{rough_norm_matrix_form},
step $(ii)$ uses the fact that $\hat{\boldsymbol{\xi}}_{l_1}$ is a generalized eigenvector ($\kappa_{l_1}$ denotes the corresponding generalized eigenvalue), as presented in Proposition \ref{proposition:eigen_regul},
and step $(iii)$ uses the definition of $V$ matrix given in \eqref{def:V_matrix},
and $c$ represents a scalar.

>>>>

	\begin{frontmatter}
		\title{ Hybrid Partial Least Squares Regression with Multiple Functional and Scalar Predictors}
		%\title{A sample article title with some additional note\thanksref{t1}}
		\runtitle{Hybrid Partial Least Squares}
		%\thankstext{T1}{A sample additional note to the title.}

@@

	\subsection{Preliminary step 1: finite-basis approximation}\label{sec: finite basis approximation}	
	Let ${b_m(t)}$ be a twice-differentiable basis of $L^2([0,1])$ whose second derivatives are also linearly independent, for example, cubic B-splines, the Fourier basis, or an orthonormal polynomial basis of degree greater than three. Using this basis, 
	the $j$th functional predictor, 
	regression coefficient, PLS component direction, and orthogonalization regression coefficient (with iteration indices suppressed) are represented as follows:
		X_{ij}(t) = \sum_{m=1}^\infty \theta_{ijm} \, b_m(t),~
		\beta_j(t) = \sum_{m=1}^\infty \eta_{jm} \, b_m(t),~
		%%%%%%%%%%%%%%%%%%%%%%
		\xi_j(t)
		\sum_{m=1}^\infty
		\gamma_{jm}
		b_m(t),
		\delta_j(t)
		\sum_{m=1}^\infty
		\pi_{jm}
		b_m(t)		
	In practice, the full set of coefficients can not be obtained  with finite sample size, as functional data are measured on a finite grid. Thus, we truncate the expansion at $M$ terms and define the truncated basis space
	\mathcal{B} := \operatorname{span}\bigl(b_1(t), \ldots, b_M(t)\bigr).
	We choose a moderately large $M$ (e.g., 15 or 20) to capture functional variation without fine-tuning, as smoothness is handled via penalization (see Section~\ref{sec: Regularization of PLS Components}).
	The truncated expansions of the predictor and coefficient are denoted as
		\widetilde{X}_{ij}(t) := \sum_{m=1}^M \theta_{ijm} \, b_m(t), \quad
		\tilde{\beta}_j(t) := \sum_{m=1}^M \eta_{jm} \, b_m(t),
	and our suggest method restricts each PLS component direction and orthogonalization regression coefficient to admit the following expansion:
	\begin{equation}\label{def:truncated_zeta_and_delta}
		\xi_j(t) = \sum_{m=1}^M \gamma_{jm} \, b_m(t),
		\delta_j(t)
		\pi_{jm}
		b_m(t)		
	This implies that all computations in this paper are carried out entirely within the subspace
	\begin{equation}\label{def:truncated_hilbert_space}
		\widetilde{ \mathcal{H} } := \mathcal{B}^K \times \mathbb{R}^p \subset \mathcal{H}.
	The $i$th hybrid predictor, projected on $\widetilde{ \mathcal{H} }$, is represented by the tuple
		\widetilde{W}_i := 
		\widetilde{X}_{i1}, \ldots, \widetilde{X}_{iK}, \boldsymbol{Z}_i
		).
	Let $\boldsymbol{\theta}_{ij}$, $\boldsymbol{\eta}_j$,  $\boldsymbol{\gamma}_j$, and
	$\boldsymbol{\pi}_j$
	denote the $M$-dimensional vectors of coefficients:
	\begin{equation}\label{def:coef_vector_form}
		\boldsymbol{\theta}_{ij} := (\theta_{ij1}, \ldots, \theta_{ijM})^\top,
		\boldsymbol{\eta}_{j} := (\eta_{j1}, \ldots, \eta_{jM})^\top ,~
		\boldsymbol{\gamma}_{j} := (\gamma_{j1}, \ldots, \gamma_{jM})^\top,~
		\boldsymbol{\pi}_{j} := (\pi_{j1}, \ldots, \pi_{jM})^\top
	For the predictors, we stack the coefficient vectors across observations into the matrix
	\begin{equation}\label{def:theta}
		\Theta_j := (\boldsymbol{\theta}_{1j}, \ldots, \boldsymbol{\theta}_{nj})^\top  \in \mathbb{R}^{n \times M},
	and construct the full design matrix 
	\begin{equation}\label{def:full_theta}
		\Theta := (\Theta_1, \ldots, \Theta_K, Z) \in \mathbb{R}^{n \times (MK + p)}.
	Let us denote the response vector as $\mathbf{y} := (y_1, \ldots, y_n)^\top$.
	Let $B, B^{\prime\prime} \in \mathbb{R}^{M \times M}$ denote the Gram matrices of the basis functions and their second derivatives, with entries
	\begin{equation}\label{def:B_matrix}
		B_{m, m'} := \int_0^1 b_m(t)\, b_{m'}(t)\, dt, \quad
		B^{\prime\prime}_{m, m'} := \int_0^1 b_m''(t)\, b_{m'}''(t)\, dt,
	for $m, m' = 1, \ldots, M$. We then define the block-diagonal matrices
	\begin{equation}\label{def:gram_block}
		\mathbb{B} := \operatorname{blkdiag}(B, \ldots, B, I_p), \quad
		\mathbb{B}^{\prime\prime} := \operatorname{blkdiag}(B^{\prime\prime}, \ldots, B^{\prime\prime}, I_p),

@@

	\subsubsection{Iterative step 2: residualization via hybrid-on-scalar regression} \label{section:sub:residualization}
	The $l$-th iteration's second step involves residualization of both predictors and responses. We first compute the individual PLS score:  
	\begin{equation}\label{def:plsscore}
		\hat{\rho}^{[l]}_i := \langle \widetilde{W}_i^{[l]}, \, \hat{\xi}^{[l]} \rangle_{\mathcal{H}},
	\end{equation}
	using the estimated PLS component direction $\widehat{\xi}^{[l]}$ obtained from Propositions \ref{proposition:eigen_regul} and \ref{proposition:linear_regul}. 
	Since   $\widetilde{W}_i^{[l]}$ are assumed to have a sample mean of zero, these PLS scores will also have a sample mean of zero.
	To obtain the $(l+1)$-th  iteration's responses and hybrid predictors, we regress the $(l)$-th iteration's responses and hybrid predictors on  
 these PLS scores by least squares and then residualize. 
	Specifically,
the $(l+1)$-th predictor is computed as a hybrid-on-scalar linear regression residual.
In the same spirit as the PLS component direction estimation step,
rather than treating the hybrid object as a long vector of concatenated function evaluations at time points and scalar vectors,
we employ a basis expansion approach to fit the entire hybrid object in one step.
Therefore, our method is computationally efficient, applicable for dense or irregular functional data, and keeps the residual smooth, as it enforces the regression coefficient, $\delta^{[l]}$, as an element of $\tilde{\mathcal{H}}$ as defined in Equation \eqref{def:truncated_hilbert_space}.
	Consequently, $\delta^{[l]}$ is obtained by minimizing a  
	least squares criterion:
	\begin{equation}\label{def:argmin_pensse}
		\hat{\delta}^{[l]} := \arg \min_{\delta \in \tilde{\mathcal{H}}} \sum_{i=1}^n 
		\| \widetilde{W}_i^{[l]} - \hat{\rho}_i^{[l]}\, \delta\|_{\mathcal{H}}^2.
	\end{equation}
  On the other hand, the $(l+1)$-th response   is computed as a scalar-on-scalar linear regression residual. 
 The following proposition demonstrates that this residualization step can be performed simply, analogous to scalar PLS.
	\begin{lemma}[Closed-form  solution]\label{proposition:closed_form_orthgonalization}
		Let us denote
		$
		\hat{\boldsymbol{\rho}}^{[l]} := (\hat{\rho}^{[l]}_1, \ldots, \hat{\rho}^{[l]}_n)^\top.
		%\quad\text{and} \quad		\bar{\rho}^{[l]} := (\hat{\boldsymbol{\rho}}^{[l]})^\top \hat{\boldsymbol{\rho}}^{[l]}.
		$
	The $(l+1)$-th iteration's  predictors and responses are   computed as:
\begin{equation}\label{algorithm_step:residualization_predictor}
	\widetilde{W}_i^{[l+1]} 
	= 
	\widetilde{W}_i^{[l]}  - 
	\frac{\widehat{\rho}_{il}}{\| 	\hat{\boldsymbol{\rho}}^{[l]}\|_2^2}
	\sum_{i=1}^n 
	\widehat{\rho}_{il}
	\widetilde{W}_i^{[l]},
	%%%%%%%%%%% 
	\quad 
	Y_i^{[l+1]} := Y_i^{[l]} - 
	\frac{
		\mathbf{y}^{[l] \top }  \hat{\boldsymbol{\rho}}^{[l]}
	}{
		\| \hat{\boldsymbol{\rho}}^{[l]} \|_2^2
	}
	\widehat{\rho}_{il}.
\end{equation} 	
	\end{lemma}
	Proof of Lemma \ref{proposition:closed_form_orthgonalization} is provided in Appendix \ref{section:proof:proposition:closed_form_orthgonalization}.
	\jongmin{proof needs to be revised}
Since $\widetilde{W}_i^{[l]}$ and $Y_i^{[l]}$ are assumed to have a sample mean of zero, their respective residuals, $\widetilde{W}_i^{[l+1]}$ and $Y_i^{[l+1]}$, also maintain a zero sample mean.

@@

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%% Support information, if any,             %%
	%% should be provided in the                %%
	%% Acknowledgements section.                %%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{acks}[Acknowledgments]
		The authors would like to thank the anonymous referees, an Associate
		Editor and the Editor for their constructive comments that improved the
		quality of this paper.
	\end{acks}

	%% if your bibliography is in bibtex format, uncomment commands:
	\bibliographystyle{imsart-nameyear} % Style BST file
	\bibliography{bibliography.bib}       % Bibliography file (usually '*.bib')

@@

		The solution to the optimization problem \eqref{def:argmin_pensse}, denoted
	$\hat{\delta}^{[l]}   \in \tilde{\mathcal{H}}$, has basis coefficients
		\hat{\boldsymbol{\pi}}_j^{[l]}
		\frac{1}{\| 	\hat{\boldsymbol{\rho}}^{[l]}\|_2^2}
		\Theta_j^{[l] \top}
		\hat{\boldsymbol{\rho}}^{[l]}
		,
	and scalar parts given by:
		\hat{\boldsymbol{\chi} }^{[l]}
		\frac{1}{\| 	\hat{\boldsymbol{\rho}}^{[l]}\|_2^2}
		\mathbf{Z}^{[l] \top}  \hat{\boldsymbol{\rho}}^{[l]}.

@@

		Now to switch our gears toward the the orthogonality. For  $l_1 > l_2 $, we have
		\begin{align*}
			\langle \widehat{\xi}_{l_1}, \widehat{\xi}_{l_2} \rangle_{\mathcal{H}, \Lambda}
			& \overset{(i)}{=}
			\hat{\boldsymbol{\xi}}_{l_1}^\top 
			\hspace{-.2em}
			(\mathbb{B} + \Lambda \mathbb{B}^{\prime \prime}) \hat{\boldsymbol{\xi}}_{l_2}
			%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
			\\	& \overset{(ii)}{=} \frac{1}{\kappa_{l_1}} (V^{[l_1]} \widehat{\boldsymbol{\xi}}_{l_1})^\top \hat{\boldsymbol{\xi}}_{l_2}
			%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
			\\	& = \frac{1}{\kappa_{l_1}}
			\widehat{\boldsymbol{\xi}}_{l_1}^\top
			V^{[l_1]} \hat{\boldsymbol{\xi}}_{l_2} 
			%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
			\\	& \overset{(iii)}{=} \frac{1}{ n^2 \kappa_{l_1}}
			\widehat{\boldsymbol{\xi}}_{l_1}^\top
			(\mathbb{B} \Theta^{[l_1]\top} \mathbf{y})(\mathbb{B} \Theta^{[l_1]\top} \mathbf{y})^\top
			\hat{\boldsymbol{\xi}}_{l_2}
			%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
			\\	& = \frac{1}{ n^2 \kappa_{l_1}}
			\widehat{\boldsymbol{\xi}}_{l_1}^\top
			(\mathbb{B} \Theta^{[l_1]\top} \mathbf{y})
			\mathbf{y}^\top
			\Theta^{[l_1] } 
			\mathbb{B}
			\hat{\boldsymbol{\xi}}_{l_2}
			%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
			\\	& = c
			\mathbf{y}^\top
			(
			\Theta^{[l_1] }
			\mathbb{B}
			\hat{\boldsymbol{\xi}}_{l_2}
			),
		\end{align*}
		where
		step $(i)$ uses \eqref{rough_norm_matrix_form},
		step $(ii)$ uses the fact that $\hat{\boldsymbol{\xi}}_{l_1}$ is a generalized eigenvector ($\kappa_{l_1}$ denotes the corresponding generalized eigenvalue), as presented in Proposition \ref{proposition:eigen_regul},
		and step $(iii)$ uses the definition of $V$ matrix given in \eqref{def:V_matrix}.
Here, $c$ represents a scalar that condenses all multiplicative terms preceding $\mathbf{y}^\top$.

		Our next task is to show that 
		$
		\Theta^{[l_1] }
		\mathbb{B}
		\hat{\boldsymbol{\xi}}_{l_2} = \mathbf{0} \in \mathbb{R}^n.
		$
		From the construction in \eqref{def:full_theta}, \eqref{def:theta} and \eqref{def:gram_block}, the
		$i$th entry of $\Theta^{[l_1] }
		\mathbb{B}
		\hat{\boldsymbol{\xi}}_{l_2}$ 
		is
		\begin{align*}
			(\Theta^{[l_1] }
			\mathbb{B}
			\hat{\boldsymbol{\xi}}_{l_2})_i
			&=
			(\theta_{i11}^{[l_1]}, \ldots, \theta_{i1M}^{[l_1]}, \ldots, \theta_{iK1}^{[l_1]}, \ldots, \theta_{iKM}^{[l_1]}, Z_{i1}^{[l_1]}, \ldots, Z_{ip}^{[l_1]})
			\mathbb{B}
			\hat{\boldsymbol{\xi}}_{l_2}
			%%%%%%%%%%%%%%%%%%%%%%%%%%%
			\\&=
			\sum_{k=1}^K
			\boldsymbol{\theta}_{ik}^{[l_1]} B
			\hat{\boldsymbol{\gamma}}_k^{[l_2]}
			+
			\mathbf{Z}_{i}^{[l_1] \top } \hat{\boldsymbol{\zeta}}^{[l_2]}
			%%%%%%%%%%%%%%%%%%%%%%%%%%%
			\\	&=
			\langle
			\widetilde{W}_i^{[l_1]}, 
			\hat{\xi}^{[l_2]}
			\rangle_{\mathcal{H}},
		\end{align*}
where the last equality uses the computation of the $i$th entry in \eqref{score_inner_product_all_n}. To expand the last quantity, we use a recursive relationship derived as follows.
\begin{align*}
	\widetilde{W}_i^{[l_1]} &= \widetilde{W}_i^{[l_1-1]} - \widehat{{\rho}}_i^{[l_1-1]} \widehat{\delta}^{[l_1-1]}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\\&\overset{(i)}{=}
	\widetilde{W}_i^{[l_1-1]}
	-
	\frac{\widehat{{\rho}}_i^{[l_1-1]}}{\| \widehat{\boldsymbol{\rho}}^{[l_1-1]} \|_2^2}
	\bigl(
	\sum_{k=1}^K {\Theta}_{k}^{[l_1-1]} 
	+
	Z^{[l_1-1]} 
	\bigr)
	\widehat{\boldsymbol{\rho}}^{[l_1-1]}
	\\&=
	\langle
	\widetilde{W}_i^{[l_1-1]}  , 
	\hat{\xi}^{[l_2]}
	\rangle_{\mathcal{H}}
	-
	\frac{\widehat{{\rho}}_i^{[l_1-1]}}{\| \widehat{\boldsymbol{\rho}}^{[l_1-1]} \|_2^2}
	\widehat{\boldsymbol{\rho}}^{[l_1-1] \top}
	\langle
	\widetilde{W}^{[l_1-1]}  , 
	\hat{\xi}^{[l_2]}
	\rangle_{\mathcal{H}}
\end{align*}
Using the regression model  we used for residualization,
		and step $(iii)$ uses 
		the closed-form formula of $\widehat{\delta}^{[l_1-1]}$ given by Lemma \ref{proposition:closed_form_orthgonalization},
		and we abuse the notation as
		$\langle
		\widetilde{W}_1^{[l_1-1]}  , 
		\hat{\xi}^{[l_2]}
		\rangle_{\mathcal{H}}:=
		\bigl(
		\langle
		\widetilde{W}_n^{[l_1-1]}  , 
		\hat{\xi}^{[l_2]}
		\rangle_{\mathcal{H}},
		\ldots,
		\langle
		\widetilde{W}^{[l_1-1]}  , 
		\hat{\xi}^{[l_2]}
		\rangle_{\mathcal{H}}
		\bigr)^\top \in \mathbb{R}^n.
		$

		Therefore, we have
		\begin{equation*}
			\Theta^{[l_1] }
			\mathbb{B}
			\hat{\boldsymbol{\xi}}_{l_2}
			=
			\langle
			\widetilde{W}^{[l_1-1]}  , 
			\hat{\xi}^{[l_2]}
			\rangle_{\mathcal{H}}
			-
			\frac{1}{\| \widehat{\boldsymbol{\rho}}^{[l_1-1]} \|_2^2}
			\widehat{\boldsymbol{\rho}}^{[l_1-1] \top}
			\langle
			\widetilde{W}^{[l_1-1]}  , 
			\hat{\xi}^{[l_2]}
			\rangle_{\mathcal{H}}
			\widehat{\boldsymbol{\rho}}^{[l_1-1]}
		\end{equation*}
		\begin{equation*}
			\hat{\boldsymbol{\pi}}_j^{[l]}
			=
			\frac{1}{\| 	\hat{\boldsymbol{\rho}}^{[l]}\|_2^2}
			\Theta_j^{[l] \top}
			\hat{\boldsymbol{\rho}}^{[l]}
			,
			~
			j = 1, \ldots, K,
		\end{equation*}
		and scalar parts given by:
		\begin{equation*}
			\hat{\boldsymbol{\chi} }^{[l]}
			=
			(\mathbf{Z}^{[l]})^\top  \hat{\boldsymbol{\rho}}^{[l]}.
		\end{equation*}
